---
title: "kluster package"
output: html_notebook
---

This is a full example of using the `kluster` package.

```{r}
# install `kluster`:
devtools::install_github("hestiri/kluster")

#loading the required packages.
if (!require("easypackages")) install.packages('easypackages', repos = "http://cran.rstudio.com/")

packages("kluster","mclust","dplyr","DT","ggplot2","tictoc","clusterGeneration","factoextra","apcluster","fpc","vegan",
         prompt = F)
#loading the data
setwd("..")
dat = read.csv(paste0(getwd(),"/data/Breast_Cancer_Wisconsin.csv"))
```

We will use `kluster::Breast_Cancer_Wisconsin` for this demo. Below is a plot of the 2 clusters in this dataset.
```{r}
ggplot(dat)+
  geom_point(aes(x=area_mean, y= texture_mean,colour=diagnosis),size=3,alph=0.3)+
  ggtitle("") +
  theme_bw() +
  theme(panel.grid.major.y = element_line(colour = "gray"),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(size=0.5, colour = "black"),
        panel.border = element_blank(), panel.background = element_blank(),
        plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text=element_text(family="Tahoma", face = "bold"),
        axis.text.x=element_text(colour="black", size = 10, face="plain"),
        axis.text.y=element_text(colour="black", size = 10, face="plain"),
        legend.position="bottom")
```

Now let's get the best estimate using the main `kluster` function -- 100 itertations of the BIC algorithm on 100 random samples.
```{r}
k = kluster(data = dat[,c("area_mean","texture_mean")],iter_klust = 100, smpl=100, algorithm = "BIC")$f_bic_k
k
```

to test other algorithm's kluster implementation results:
```{r, warning= FALSE, echo= FALSE}
eval= data.frame(kluster_eval(data = dat[,c("area_mean","texture_mean")], clusters = 2, iter_sim = 2, iter_klust = 7, smpl = 100)$sim)
# eval
```
```{r}
datatable(eval, options = list(pageLength = 8), filter = 'bottom')
```



Now we can test results of kluster across with the implementation of original algorithms. Here is an example of the BIC algorithm:
```{r, warning= FALSE, echo= FALSE}
sim.bic = kluster_sim(data = dat[,c("area_mean","texture_mean")], clusters = 2, iter_sim = 2, iter_klust = 7, smpl = 100, algorithm = "BIC")$sim
# sim.bic
```

```{r}
datatable(sim.bic, options = list(pageLength = 3), filter = 'bottom')

```

